Repeatability of Pre-Flashover Fire Patterns on Gypsum Wall Board

### 2015-03-23 Projects/Madrzykowski_Thesis/Report/Madrzykowski_Thesis.tex
Working on Ch 6 plus  Wall Construction Table needs help

The results are presented for each of the four burner positions, (0.61~m, 0.31~m, 0.15~m and 0.0~m), and each of the three fuels in Appendix X.  For the natural gas fueled experiments only two replicates were conducted at each position given the high level of control and repeatability.  For the gasoline and polyurethane fueled experiments, three replicates were conducted at each position. The position is measured as the distance between the edge of the burner closest to the wall and the wall.  The results of the wall temperatures and heat fluxes are presented as averages of the replicate experiments.  In the sections below, examples of the averaged data from the gasoline fueled experiments will be presented to demonstrate the impact on distance on the conditions measured on the wall. Then a composite graph of the measured thermal impacts on the wall versus distance will be shown for each fuel type.

@@

Figure~\ref{IW_SC_TWGAS} shows the temperatures of the array of thermocouples located along the vertical centerline of the gasoline fueled burner.  The beads of the thermocouples were installed in the wall and were flush with the wall surface. In each graph, the average temperature of the replicate experiments, for each thermocouple position is shown.  Moving the burner from

@@

Figure~\ref{IW_SC_TWGAS} relabelshows the temperatures of the array of thermocouples located along the vertical centerline of the gasoline fueled burner.  The beads of the thermocouples were installed in the wall and were flush with the wall surface. In each graph, the average temperature of the replicate experiments, for each thermocouple position is shown.  Moving the burner from

@@

 Impact of Wall Construction
 Additional experiments were conducted with the natural gas and gasoline fires to examine the impact of the method of construction of the walls on the development of the fire patterns.  Three cases were examined: 1) Open back, 12.7~mm thick, regular core, painted gypsum board on the front of a wood frame only, 2) Closed back, 12.7~mm thick, regular core, painted gypsum board on the front and back of a wood frame representative of an interior wall, and 3)	Insulated, 12.7~mm thick, regular core, painted gypsum board on the front and back of a wood frame, with fiberglas batt insulation in the voids of the wood frame representative of an exterior wall.

The fuel or burner was centered on the middle panel of the rear wall.  As in the fire source characterization experiments the top of the fuel in each case was positioned 7.5~cm above the floor of the compartment.  Each of the painted gypsum board panels that were used for a fire pattern had a 5 cm grid drawn on it to assist with the determination of the geometry of the fire pattern.

\subsection{Experimental Arrangement}
\subsection{Results}

@@

>>>>

The results are presented for each of the four burner positions, (0.61~m, 0.31~m, 0.15~m and 0.0~m), and each of the three fuels in Appendix X.  For the natural gas fueled experiments only two replicates were conducted at each position given the higher level of control and repeatability.  For the gasoline and polyurethane fueled experiments, three replicates were conducted at each position. The position is measured as the distance between the edge of the burner closest to the wall and the wall.  The results of the wall temperatures and heat fluxes are presented as averages of the replicate experiments.  In the sections below, examples of the averaged data from the gasoline fueled experiments will be presented to demonstrate the impact on distance on the conditions measured on the wall. Then a composite graph of the measured thermal impacts on the wall versus distance will be shown for each fuel type.

@@

Figure~\ref{IW_SC_TWGAS} shows the temperatures of the array of thermocouples located along the vertical centerline of the gasoline fueled burner.  The beads of the thermocouples were installed in the wall and were flush with the wall surface. In each graph, the average temperature of the replicate experiments, for each thermocouple position is shown.  Moving from 0.2 m above the top surface of the burner up to 2.2 m above the top of the burner surface in increments of 0.2 m.

@@

Figure~\ref{IW_HFC_TWGAS} shows the temperatures of the array of total heat flux gauges located along the vertical centerline of the gasoline fueled burner.  The sensing surface of the gauges were installed flush with the wall surface and exposed to the fire. In each graph, the average temperature of the replicate experiments, for each total heat flux position is shown.  Moving from 0.2 m above the top surface of the burner up to 1.2 m above the top of the burner surface in increments of 0.2 m.

\begin{figure}[p]
	\centering
	\includegraphics[width=2.5in]{../Figures/HF_Center_TWGas_0_61m_RepsAvg}
	\includegraphics[width=2.5in]{../Figures/HF_Center_TWGas_0_31m_RepsAvg}\\
	\includegraphics[width=2.5in]{../Figures/HF_Center_TWGas_0_15m_RepsAvg}
	\includegraphics[width=2.5in]{../Figures/HF_Center_TWGas_0m_RepsAvg}\\	
	\caption[Wall heat fluxes along the centerline of the gasoline fueled burner]{Wall heat fluxes along the centerline of the gasoline fueled burner.  The heat fluxes are averages of three replicate experiments. Each graph displays data from a different distance between the edge of the burner and the instrumented wall.  Beginning with the top left position and moving from left to right the distances are 0.61~m, 0.31~m,0.15~m,and 0~m.}
	\label{IW_HFC_TWGAS}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{../Figures/nctw_ng_heatflux}\\
	\caption[Summary of wall heat fluxes along the centerline of the natural gas fueled burner]{Summary of the heat flux to the non-combustible wall along the verticle centerline of the natural gas fueled burner compared to the position away from the wall.}
	\label{nctw_ng_heatflux}
\end{figure}

@@

 Impact of Wall Construction

 The initial compartment experiments were conducted to examine the impact of the method of construction of the walls on the development of the fire patterns. This data determined how the walls for the remaining compartment fire pattern experiments would be constructed.  Three cases were examined: 1) Open back, 12.7~mm thick, regular core, painted gypsum board on the front of a wood frame only, 2) Closed back, 12.7~mm thick, regular core, painted gypsum board on the front and back of a wood frame representative of an interior wall, and 3)	Insulated, 12.7~mm thick, regular core, painted gypsum board on the front and back of a wood frame, with fiberglas batt insulation in the voids of the wood frame representative of an exterior wall. From the instrumented gypsum wallboard experiments, the peak temperatures on the backside of the wallboard occured when the burners were positioned against the wall.  The peak temperatures were similar for all of the fuels,approximately 80~$^\circ$C, although the duration of the higher temperatures on the rear of the wallboard was longer for the natural gas and the gasoline fueled burners. Therefore the wall construction experiments were only conducted with the natural gas and gasoline fueled burners as the impact of insulating the back of the walls exposed to the fuels with the highest energy would produce the most significant differences, if any, to the fire patterns.

The rear wall of the compartment was constructed in the same manner for a given set of experiments. The burner was centered on the middle panel of the rear wall.  As in the fire source characterization experiments the top of the fuel in each case was positioned 7.5~cm above the floor of the compartment.  Each of the painted gypsum board panels that were used for a fire pattern had a 5 cm grid drawn on it to assist with the determination of the geometry of the fire pattern.

\begin{table}
	\centering
	\begin{tabular}{|c|c|c|c|c|c|} {\textwidth}
		\hline
	Fuel (number of tests) (\%)   &       Wall Type	            &   Max. Height (m)	            & Max. Width (m)    &  Height at Max. Width (m)    &  Area (m $^2$)  \\ \hline
		Natural Gas (10)    &   Open 		& 0.74 	$\pm$ 0.12 (16~\%)   &	0.24 $\pm$ 0.06 (25~\%) 	& 0.41 $\pm$ 0.07 (17~\%)	& 0.15 $\pm$ 0.05 (33~\%) 	\\
		Natural Gas (10)    &   Closed 		& 0.74	$\pm$ 0.06 (8~\%)    &	0.24 $\pm$ 0.02 (7~\%) 		& 0.43 $\pm$ 0.03 (7~\%)  	& 0.14 $\pm$ 0.01 (6~\%) 	\\
		Natural Gas (10)    &   Insulated 	& 0.71	$\pm$ 0.10 (14~\%)   &	0.23 $\pm$ 0.04 (17~\%) 	& 0.42 $\pm$ 0.07 (17~\%) 	& 0.15 $\pm$ 0.02 (14~\%)  	\\
		Gasoline (12)       &   Open    	& 0.83	$\pm$ 0.15 (18~\%)   &	0.28 $\pm$ 0.09 (32~\%) 	& 0.44 $\pm$ 0.14 (32~\%) 	& 0.17 $\pm$ 0.04 (24~\%)  	\\
		Gasoline (10)       &   Closed  	& 0.80	$\pm$ 0.10 (13~\%)   &	0.29 $\pm$ 0.04 (14~\%) 	& 0.44 $\pm$ 0.06 (14~\%) 	& 0.17 $\pm$ 0.04 (24~\%)  	\\
		Gasoline (10)       &   Insulated	& 0.82 	$\pm$ 0.09 (11~\%)   &	0.25 $\pm$ 0.08 (32~\%) 	& 0.40 $\pm$ 0.13 (32~\%) 	& 0.13 $\pm$ 0.04 (30~\%)  	\\

		\hline
	\end{tabular}
	\caption[Comparison of fire pattern dimensions as a function of wall construction type]{Comparison of fire pattern dimensions as a function of wall construction type. Values presented with 95\% confidence limits.}
	\label{tab:Fire_Pattern_Dimensions_Wall_Construction}
\end{table}

@@

Fire Dynamics Simulator~\cite{FDS_Users_Guide} is a computational fluid dynamics (CFD) model developed and maintained by NIST that solves a form of the Navier-Stokes equations appropriate for low-speed, thermally driven flow with an emphasis on smoke and heat transport from fires. Within a CFD model, the room or building is divided into small three-dimensional rectangular control volumes or computational cells. The cells are contained together within a larger volume known as the computational domain. The CFD model computes the density, velocity, temperature, pressure, and gas concentrations in each cell. Based on the laws of conservation of mass, momentum, and energy, the model tracks the generation and movement of fire gases. One of the most important aspects of FDS is that it is mathematically verified~\cite{FDS_Verification_Guide}and validated against fire test data~\cite{FDS_Validation_Guide}to ensure that it is accurate and provides the expected results, given appropriate input data. A complete description of the FDS model is provided in the FDS Technical Reference Guide~\cite{FDS_Tech_Guide}.

Smokeview is a software tool designed to visualize simulation results from FDS~\cite{Smokeview_Users_Guide}. Smoke-view visualizes smoke and other attributes of the fire simulation by showing tracer particle flow and two dimensional (2D) or three dimensional (3D) shaded contours of gas flow data, including temperature and flow vectors showing flow direction and magnitude. Smokeview has the ability to visualize fire and smoke by displaying a series of partially transparent planes where the transparencies in each plane (at each grid node) are determined from soot densities computed by FDS. Smokeview can also visualize data at particular snapshots in time using 2D or 3D contours of data that show temperature, flow direction, and flow magnitude.

Input data was collected from the experimental results presented in Chapters XXs and from literature sources in order to simulate the experiments      a fire using FDS or any other fire model. For the simulation results presented in this document, information was obtained from two primary sources
